---
title: "Different Aspects of Abstractness Predict Vocabulary Acquisition Across Languages"
shorttitle: "Crossling Abstractness"
author:
  - name          : "Erin Campbell"
    affiliation   : "1"
    corresponding : yes    
    email         : "eecamp@bu.edu"
    address       : "2 Silber Way, Boston, MA 02215"
  - name          : "Charles P. Davis"
    affiliation   : "2"
  - name          : "Amy Lieberman"
    affiliation   : "1"
  - name          : "Naomi Caselli"
    affiliation   : "1"


affiliation:
  - id            : "2"
    institution   : "Department of Psychology & Neuroscience, Duke University, Durham, NC"
  - id            : "1"
    institution   : "Wheelock College of Education & Human Development, Boston University, Boston, MA"


bibliography: ["crossling_abstractness_citations.bib"]

lineno: true
figsintext: yes

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
latex_engine      : xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

# trackdown::update_file(file="crossling_abstractness_manuscript.Rmd",gpath="trackdown/CrosslingAbstractness",hide_code=TRUE)
```

```{r libraries}
library(tidyverse)
library(glue)
library(cowplot)
```

```{r source-and-load}
source("analysis/summarize_effects.R")
source("analysis/clusters_and_correlations.R")

n_participants <- read_rds("models/n_participants.rds")
```

```{r reporting-functions}
print_effect_size_summary <- function(variable_name) {
  summary_data <- summaries_summary %>%
    filter(variable == variable_name)
  
  if (nrow(summary_data) == 0) {
    cat("Variable not found.")
    return(NULL)
  }
  
  mean_effect_size <- round(summary_data$mean_effect_size,2)
  min_effect_size <- round(summary_data$min_effect_size,2)
  max_effect_size <- round(summary_data$max_effect_size,2)
  
  glue("B = ", {mean_effect_size}, " (range: ", {min_effect_size}, " - ", {max_effect_size}, ")\n", sep = "")
}
```

# Abstract

of vocabulary data from `r n_participants` children

<!-- # where to send this? -->
<!-- Open Mind, CogSci (the journal, not the conference), Cognition, DevSci, Language and Cognition -->

# Introduction

Young children rapidly learn new words, but some words are slower to enter children's vocabularies than others. For example, infrequent words are challenging for children to learn [CITE], as the child has few exposures from which to draw a robust mapping between the word and the referent [CITE]. Abstract words are the quintessential "hard words" [@gleitman2005], but the explanatory mechanism is murkier, in part because abstractness encompasses several interrelated lexical properties. In a recent consensus paper, abstractness was defined as:
> "(1) (historical): Referring to the quality of a concept (or word meaning) that has no sensory or motor salience (in opposition to concrete) in that it cannot be seen, heard, touched, felt, smelled, tasted or acted upon. (2) (contemporary): The quality of a concept (or word) whose meaning is understood primarily on the basis of language, but also draws from interoceptive experiences, including emotion, introspection, and metacognition. Abstract concepts are often exemplified by perceptually dissimilar associated actions and events." [@reilly2024].

Within this single definition, we have multiple, non-exclusive, potential challenges for word learning.
1) Abstract words can be characterized by an absence of sensory or motor experience with the referent [e.g., want, @paivio1971; @paivio1991]. 
2) Relatedly, words *with* sensory or motor salience can bear perceptual resemblance to their meanings [*iconicity*, e.g., the English word "moo" refers to and resembles the sound a cow makes], particularly in sign languages, several Niger-Congo languages, Quechua, Japanese, Korean, Turkish, and Basque, among others [@akita2019]. Abstract words afford less opportunity for iconic mappings between the wordform and the referent, so the phonological form of the word is less informative for its meaning meaning.
3) Abstract words may involve interoception, emotion, introspection, metacognition, or the characterization of relationships between entities. 
4) Abstract words may be less contextually specific [e.g., problem, @schwanenflugel1983; @hoffman2013; @davis2020], and so are exemplified by " perceptually dissimilar associated actions and events." [@reilly2024]. For example, unlike words like "ball" or "dog", there is no single situation that clearly teaches what "thing" means[^1].

[^1]: Philadelphian readers who have tried to explain "jawn" to an out-of-towner are likely familiar with this particular challenge.

It is fairly uncontroversial to state that children learn abstract words later [@bergelson2013; @dellarosa2010; @montefinese2019; @ponari2018; @ramey2013; @reilly2007], but why exactly they learn abstract words later is open to debate. The most oft-cited reason why children tend to acquire abstract words later than concrete ones is because abstract words lack a clear sensory-perceptual referent, and thus are characterized by referential uncertainty [e.g., @bergelson2013]. However, this explanation rests on a negative definition of abstract words---namely, whereas concrete words refer to things that can be experienced directly through the senses, abstract words refer to things that cannot, and must be instead defined by other words [@brysbaert2014]. This view on abstract words is increasingly coming into question, with a growing literature pointing to the richness of experiences associated with abstract words, emphasizing not only that understanding abstract words tends to rely more on language associations [@paivio1991; @borghi2019] and that they tend to be more ambiguous and deployed in diverse lexical and situational contexts [@schwanenflugel1983; @hoffman2013; @davis2020; @barsalou2018] but also that they tend to be highly associated with emotional experience [@vigliocco2013] and social knowledge [@wilson-mendenhall2013]. There are also differences between abstract and concrete words in iconicity (i.e., the degree to which a word's meaning is reflected in its form [@hinojosa2021; @winter2023], such that abstract words counterintuitively tend to be rated as more iconic. Assuming a multidimensional perspective on abstract words---that is, that they have a range of experiential properties beyond just a lack of sensory referents---allows us to unpack what properties of abstractness lead to differences in word learning. In the following sections, we consider four different ways in which a word meaning might be abstract: a lack of associated perceptual experience; social, emotional, or interoceptive meaning; iconic vs. arbitrary mappings between word form and meaning, and the ability to use a word in a variety of distinct contexts.

## Different Constructions of Abstractness

### Abstractness as absent of sensorimotor experience

Embodied theories propose that sensorimotor processes are essential for learning the meanings of words (e.g., CITE). Under such hypotheses, words with referents that afford greater sensorimotor experience should be easier to acquire. Greater sensorimotor experience could be thought of both in terms of the saliency of the experience and in terms of the multisensoriness of the experience, and both operationalizations have attested links to processing and learning. The words in children's vocabularies seem to reflect their individual perceptual experiences; for example, English-speaking sighted children, but not blind children, were more likely to produce words when the words are more strongly associated with visual experience [@campbell2023]. @seidl2023 observed that English words tend to be learned earlier when their semantic associations span multiple senses. In related experimental work, @seidl2023 observed that 2-year-olds learn words more readily when they are able to interact with the referents than when they are only able to see the referents.

In English, body-object-interaction ratings predict words' AoA and children's lexical processing [@suggate2017; @wellsby2014; @inkster2016], such that words with high sensorimotor interaction ratings are acquired earlier and processed faster [@thill2016; @pexman2019]. The motor aspect of this relationship may offer additional explanatory power beyond the sensory association effects described in previous paragraphs: the processing speed of high body-object-interaction words (but not low-BOI words) is linked to children's fine motor skills [@suggate2017]. Work with adults finds that pairing novel words with motor actions related to their meaning improves learning [@casasanto2019; @vukovic2019], and that activation of the motor cortex is involved in the acquisition of novel words [@luizzi2010; @vukovic2019]. This effect may be particularly subject to cross-linguistic variation, given the wide cross-cultural variation in infant's motor development [@cintas2009; @lohaus2014].

### Abstractness as internal/relational/social-emotional

"Abstract meanings are typically associated with more emotional/affective experience (Kousta et al., 2011; Lund et al., 2019; Newcombe et al., 2012; Ponari et al., 2018; Vigliocco et al., 2014), although not all abstract words are affect-laden. Similarly, abstract concepts, particularly emotional ones, are rated as evoking more inner and interoceptive experiences than concrete concepts (Connell et al., 2018; Kelly et al., in press; Lynott et al., 2020; Villani et al., 2021). In addition to language and emotion, abstract concepts are related to visual or motor experience, social constellations, and mental states (Harpaintner et al., 2020; Kiefer et al., 2022; Trumpp et al., 2024; Ulrich et al., 2022)."

### Abstractness as wordform arbitrariness

Arbitrariness to be a requisite property of language. Language is language because words are abstract symbols of their meaning. Words with an iconic (or motivated) mapping between form and meaning have originally been relegated to the fringes of linguistic discussion. Work on onomatopoeias, ideophones, and systematicity show that iconicity is a more prevalent phenomenon than previously recognized []. Across languages, words that are higher in iconicity tend to be learned earlier [@nielson2021; @sidhu2021; @perry2015]. Some speculate that this is due to easier mapping between the referent and the wordform. Others, however, find this explanation unsatisfying: do young children, who have limited experience, have the world knowledge necessary to recognize these perceptual mappings? It could be that effects of iconicity on acquisition are related to phonological constraints on the meaning of highly-iconic words: highly-iconic words are likely to carry perceptual information that aligns with the modality of the language [e.g., iconic spoken language words often pertain to auditory referents; iconic sign language words often depict auditory or tactile information; @winter2023; @thompson2020]. Further, despite these modality-specific affordances, "abstract" words tend to be rated as *more* iconic [@hinojosa2021; @winter2023] *(by conjecture perhaps because abstract words are sometimes associated with emotion meanings? and perhaps emotional meanings are rated as more iconic)*.

### Word Meaning Abstract from Specific Circumstance

The diversity of contexts a word appears in predicts word learning in both children [@hills2010; @rosa2022; @kachergis2009] and adults [@johns2015]. These effects have been interpreted in line with two processes in early word learning: the lure of the associates, which suggests that words learned earlier tend to be better connected with a range of other known words  [@hills2009]; and preferential acquisition, where words learned earlier tend to be used in a range of distinct contexts [@hills2010]. The direction of these context diversity effects are variable; in some contexts, verbs(?), are more likely to be learned when the context is more consistent, whereas nouns(?) with diverse contexts are more likely to be learned. More contextually diverse words also tend to be more abstract [@hoffman2013], as abstract words like *idea* have more flexible meanings that can be deployed in a range of contexts, as compared to concrete words like *spaceship*.

## Developmental Change in Mechanisms

It's possible that the mechanisms supporting acquisition of abstract words change across development. As children mature, their growing cognitive, social, and linguistic skills may pave the way for increasingly sophisticated word learning [@meylan2022]. Particularly if we conceptualize abstract words as acquired primarily through linguistic context, as children's linguistic comprehension continues to mature, we should expect to find that effects of linguistic context on AoA strengthen across age (and perhaps ultimately weaken, as children reach ceiling on their ability to extract meaning from linguistic context). Similarly, if word learning is supported by sensorimotor experience with a word referent, as children's motor abilities advance, and children are increasingly able to manipulate objects and independently locomote, we would expect to find xxx [cite].

## Crosslinguistic Variability in Mechanisms

The age of acquisition of words is correlated to varying degrees across languages [@luniewska2016; @luniewska2019].
Studying patterns of language acquisition across languages is informative for understanding which acquisition mechanisms are persistent across languages and cultures, versus which mechanisms are specific to a particular developmental or linguistic context. For example, work with Mandarin learners called into question the universality of the *noun bias* in children's early vocabulary [@bornstein2004; @tardif2003; @kauschke2007; @hardegger2024 cite]. The strength and direction of lexical effects on age of acquisition can vary across languages; for instance, while frequency and shorter word length consistently predict earlier word production across many languages, effects of emotional valence are more variable [@braginsky2019]. Additionally, within a given language, there are commonly strong correlations between lexical properties; e.g., in English, xxx is strongly correlated with xxx [cite]. Thus, it can be difficult to interpret whether an early-learned word is early-learned because of *Lexical Property X* or *Lexical Property Y*. By looking across languages, we can utilize cross-linguistic variability in the distribution of lexical properties to achieve better power to isolate individual lexical properties.

## The Present Study

The present study examines the cross-linguistic stability of associations between lexical properties traditionally linked to abstractness and age of acquisition. In doing so, we hope to uncover, across many languages: what is it about abstractness that makes a word challenging to learn? This study is only possible due to massive effort from researchers around the world to document lexical properties and vocabulary. We leverage lexical ratings collected by hundreds of researchers from thousands of participants. The vocabulary data presented here span tens of thousands of young children. Today's backdrop of open science practices makes such an analysis feasible.

# Methods

## CDI

the MacArthur-Bates Communicative Development Inventory (CDI), a vocabulary measure in which parents are asked to indicate whether their child understands and/or produces each of several hundred words @fenson1994. Since its conception, the CDI has been adapted into many languages, validated with many populations, and administered to many thousands of children.

When the CDI is adapted into a new language, researchers of each language tend to consult previous CDIs, parents, educators, and child development experts to curate a culturally-relevant set of early-learned words [@jaruskova2023]. Thus, the CDIs contain partially-overlapping subsets of words (e.g., words for mother and father are ubiquitous across CDI forms, but acorn, magpie, and paprika appear on only the Latvian, Beijing Mandarin, and Hungarian forms, respectively).

We pull data from Wordbank, a large, open database of CDI data [@frank2017], using the wordbankr R package [@braginsky2024]. For the present analyses, we take the full set of participants from Wordbank as of `r date()`. This results in a sample of N = `r n_participants`.

## Lexical Properties

All lexical ratings are first re-scaled (within languages) to a 1 to 10 scale for better comparability across languages and lexical properties. For zipfian-distributed norms (e.g., corpus-based frequency), norms were first logged and then scaled.

To achieve maximum lexical rating coverage across words and languages, while recognizing that language-specific ratings by native speakers are likely the best source of lexical data, we developed the following rules for imputation:

-   For norms pertaining to the underlying concept (e.g., concreteness, perceptual strength, emotional valence):

    -   If norms are available for a specific word in a specific language, the scaled value will be used.

    -   If norms are not available for a specific word or not available in a specific language, then the mean rating for the words' translation equivalent from all available languages will be used.

    -   If norms are not available for a specific word or its translation equivalent, then any words without norms will be excluded from the analysis.

-   For norms pertaining to the wordform (e.g., frequency, iconicity, phonological properties):

    -   If norms are available for a specific word in a specific language, the scaled value will be used.

    -   If norms are not available for a specific word but are for other words in that language, then any words without norms will be excluded from the analysis.

    -   If there are no norms at all for a language along one of these scales, then any set of analyses using that norming scale will exclude that language.

<https://docs.google.com/spreadsheets/d/1fivS85qayBJtp_nXjVmAvwZn7u6F517RFur0UPTJcAw/edit#gid=0>

#### Imageability

#### Perceptual properties

More recently, researchers have developed more specific measures of words' perceptual associations. Beyond general perceptibility, these norms separate associations by modality (visual, tactile, auditory, haptic, gustatory, interoceptive), yielding the modality specific ratings, an overall strength rating, and a measure of the spread of ratings across modalities [@lynott2020]. Such specificity allows researchers to test whether information from certain senses, like vision or touch, may be privileged over others [cite], as well as whether receiving sensory information from multiple modalities may support acquisition [e.g., @seidl2023].For example, the Lancaster Sensorimotor Norms, created for English , Using similar instructions, sensory norms have been created for other languages, including Chinese [@zhong2022], Spanish [@diezalamo2017], French [@miceli2021], Russian [@miklashevsky2018], and Italian [@vergallito2020].

#### BOI

Body-object interaction (BOI) ratings measure the extent to which a human body can physically interact with a word's referent [@siakaluk2008], and in particular, seem to index ease of graspability [@heard2019]. In neuroimaging studies, semantic processing of high BOI (vs low BOI) words is associated with neural regions involved in grasping (Hargreaves et al., 2012). In line with embodied theories of cognition, words' BOI ratings predict the speed of lexical processing in children and adults [CITE] as well as age of acquisition [@thill2016; @pexman2019; @muraki2022]. Child-centered BOI ratings, gathered by asking parents of elementary-aged children "how easily the average 6-year-old can physically interact (using the body: hands, mouth, etc.) with what each word represents" [@muraki2022, pg.9], may capture this variability even better; though in English, child BOI ratings and adult BOI ratings correlate strongly (*r* = 0.67) with each other [@muraki2022].

#### Iconicity

At the time of publication, iconicity norms were only available for English [@winter2023], ASL [@caselli2017; @sehyr2021], BSL, Japanese [CITE], and Spanish [@hinojosa2021]. We did not attempt to interpolate values for this property given its intrinsic connection to phonological form, which varies cross-linguistically.

#### xxx

```{r}
CDI_mega_word_list_with_ratings <- read_rds("norms/CDI_mega_word_list_with_ratings.rds")

CDI_mega_word_list_with_ratings_distinct_lemmas <- CDI_mega_word_list_with_ratings %>% dplyr::select(uni_lemma, contains("average"), english_iconicity_rating) %>% distinct(uni_lemma, .keep_all = TRUE) 
rating_cols <- names(CDI_mega_word_list_with_ratings_distinct_lemmas) %>%
  setdiff(c("uni_lemma")) %>%
  keep(~ is.numeric(CDI_mega_word_list_with_ratings_distinct_lemmas[[.]]))
top_bottom_words <- map_dfr(rating_cols, function(rating_col) {
  df <- CDI_mega_word_list_with_ratings_distinct_lemmas %>%
    filter(!is.na(.data[[rating_col]])) %>%
    arrange(desc(.data[[rating_col]]))

  total_lemmas <- nrow(CDI_mega_word_list_with_ratings_distinct_lemmas)
  non_na_lemmas <- nrow(df)
  coverage <- glue({round(non_na_lemmas / total_lemmas * 100, 1)},"%")

  top_3 <- df %>%
    slice(1:3) %>%
    mutate(label = str_c(uni_lemma, " (", round(.data[[rating_col]], 2), ")")) %>%
    pull(label) %>%
    str_c(collapse = ", ")

  bottom_3 <- df %>%
    slice_tail(n = 3) %>%
    mutate(label = str_c(uni_lemma, " (", round(.data[[rating_col]], 2), ")")) %>%
    pull(label) %>%
    str_c(collapse = ", ")

  tibble(
    Property = rating_col,
    top_3 = top_3,
    bottom_3 = bottom_3,
    uni_lemma_coverage = coverage
  )
})

top_bottom_words %>% View()

property_sources <- tibble::tribble(
  ~Property,                        ~PrettyName,              ~Source,
  "average_visual_rating",         "Visual",                 "Dutch [@speed2022], English [@lynott2020], Estonian [], French [@miceli2021], Italian [@vergallito2020], Mandarin [@zhong2022], Russian [@miklashevsky2018], Spanish [@diez-alamo2017]",
  "average_auditory_rating",       "Auditory",               "Dutch [@speed2022], Estonian, French [@miceli2021], Italian [@vergallito2020], Mandarin [@zhong2022], Russian [@miklashevsky2018], Spanish",
  "average_haptic_rating",         "Haptic",                 "Dutch [@speed2022], Estonian, French [@miceli2021], Italian [@vergallito2020], Mandarin [@zhong2022], Russian [@miklashevsky2018], Spanish",
  "average_olfactory_rating",      "Olfactory",              "Dutch [@speed2022], Estonian, French [@miceli2021], Italian [@vergallito2020], Mandarin [@zhong2022], Russian [@miklashevsky2018], Spanish",
  "average_gustatory_rating",      "Gustatory",              "Dutch [@speed2022], Estonian, French [@miceli2021], Italian [@vergallito2020], Mandarin [@zhong2022], Russian [@miklashevsky2018], Spanish",
  "average_interoceptive_rating",  "Interoceptive",          "Dutch [@speed2022], English [@lynott2020], French?? [@miceli2021], Mandarin [@zhong2022]",
  "average_emotionalarousal_rating", "Emotional Arousal",    "Dutch [@verheyen2020], English [@mohammad2018], Estonian, French [CITE], Italian [], Mandarin [@xu2022], Swedish [@blomberg2015], Turkish [@torkamani-azar2019; @conca2024]",
  "average_imageability_rating",   "Imageability",           "Croatian [@peti-stantic2021], Dutch [@verheyen2020], French [@desrocher2009], Italian, Mandarin [@su2023], Norwegian, Portuguese [@soares2017], Russian [@miklashevsky2018], Spanish, Swedish [@blomberg2015], Turkish [@conca2024]",
  "average_concreteness_rating",   "Concreteness",           "Croatian [@peti-stantic2021], Dutch [@verheyen2020], English [@brysbaert2014], Estonian [], French [@bonin2018], Italian, Portuguese [@soares2017]",
  "average_CD_rating",             "Context Diversity",      "*all* [Wikipedia], except ASL, BSL, etc.",
  "average_boi_rating",            "Body Object Interaction","English [@body], French [@miceli2021], Spanish []",
  "english_iconicity_rating",      "Iconicity",              "ASL [@caselli2017a], BSL [@ortega2025], English [@winter2023a], Spanish [@hinojosa2020], Japanese [CITE]",
  "average_maxperceptual_rating",  "Max Perceptual",         "Derived from Lynott et al. norms",
  "average_freq_rating",           "Word Frequency",         "*all* [Wikipedia], except ASL, BSL, etc.",
  "average_socialness_rating",     "Socialness",             "N/A (placeholder, all NA)",
  "average_emotionalvalence_rating", "Emotional Valence",    "Not in this dataset")

properties_summary <- left_join(property_sources, top_bottom_words)%>% mutate(Property=PrettyName) %>%
  select(-PrettyName)
  kableExtra::kable()
  
```

### Covariates

All models included the following variables as covariates: age (from Wordbank, rounded to the nearest month), frequency (from Wikipedia, or from subjective native speaker ratings), phonological complexity (for spoken languages, word length; for sign languages, xxx), lexical category (from Wordbank; levels included: nouns, predicates, function words, other)

```{r ratings-table}
# language by property, with citation in cell
```

# Results

## Relationships among Abstractness Variables

First, we computed pairwise Pearson's correlation coefficients for all variables, displayed in Figure 2, together with histograms showing the distribution of each variable and scatterplots showing how variables relate to each other. Here, we summarize across languages, but the by-language plots can be found in Supplementals (possibly as confusion matrix). Also the correlations for a given property across languages in supplement

<!-- stats? maybe for this section, confusion matrix is enough -->

```{r variable-relationships, fig.height=7}

corrplot(average_ratings_matrix, "color",type="upper", order = "alphabet",
         tl.col="black", addCoef.col = "white", col=colorRampPalette(c("#CC0000","white","#3B9AB2"))(200))
```

```{r variable-relationships-dendro, fig.height=3, fig.width=7}
average_ratings_dendro
```

## Abstractness effects on word production

First, we fit a baseline statistical models for each language where word production is modeled as a combination of age, frequency, and phonological complexity. We then added each of our target predictors. For each target predictor for each language, it is thus possible to extract an effect size estimate of the target predictor and its interaction with age on the likelihood of word production. To explore heterogeneity and to estimate central tendency and variation in the effects, we borrow from meta-analytic techniques, treating each language as a study, and creating a weighted average of the effect size.

Figure \@ref(fig:summary-of-maineffects) shows the coefficient estimate for each predictor in each language; see Table xxx for a full accounting of abstractness effect sizes. We find that iconicity is the strongest predictor of acquisition (mean across languages and measures: `r print_effect_size_summary("iconicity")`). Many of the sensorimotor variables were relatively strong predictors of word production, including body-object-interaction (`r print_effect_size_summary("Body Object Interaction")`), imageability (`r print_effect_size_summary("imageability")`), hapticness (`r print_effect_size_summary("haptic")`), and visualness (`r print_effect_size_summary("visual")`). The effects of the social-emotional variables socialness (`r print_effect_size_summary("socialness")`) and emotional arousal (`r print_effect_size_summary("emotionalarousal")`)--as well as several of the other sensorimotor variables (olfactoriness, interoceptiveness, auditoriness, and gustatoriness)--had much smaller mean effects on children's word production. Interestingly, concreteness was not reliably associated with word production across languages (`r print_effect_size_summary("Concreteness")`). Finally, the effect of context diversity was highly variable, with some languages showing large positive effects (e.g., ), while others showed large negative effects. Looking across languages, words which were used in more contexts were less likely to be produced (`r print_effect_size_summary("contextdiversity")`).

```{r summary-of-maineffects, fig.cap="Effect size estimates (from logistic regression) of each of the abstractness variables studied. Each dot represents the effect size of one language, with filled-in dots showing significant effects and empty dots indicating that the effect was not significant in that language. Black dot and whiskers show mean effect size with standard error."}
summary_of_maineffects
```

### Developmental Change

Positive age interactions can be seen in at least xxx out of xxx languages for visualness, hapticness, body-object-interaction, and imageability. Conversely, there are negative age interactions for auditoryiness and iconicity in at least xxx out of xxx languages. This suggests that certain sensorimotor properties (chiefly, children's ability to see and touch a word's referent) facilitate learning more so later in development, while iconicity and auditoriness[^1] facilitate learning earlier in development. This result is consistent with the speculation above that xxx.

[^1]: The interaction between auditoriness and age was near 0 for both American Sign Language (B = xxx) and British Sign Language (B = xxx).

```{r summary-of-interactions, fig.cap="Effect size estimates (from logistic regression) of the interaction between each of the abstractness variables and age. Each dot represents the effect size of one language, with filled-in dots showing significant interactions and empty dots indicating that the effect was not significant in that language. Black dot and whiskers show mean effect size with standard error. Positive values show that effects are stronger for older children, while negative values show that effects are larger for younger children."}
summary_of_interactions
```

# What drives abstractness effects?

Finally, we conduct an exploratory analysis aimed at measuring which of these abstractness variables best accounts for variability in words' production. Because these variables are interrelated, as shown in Figure xxx, when adding them to the same model, there is a risk of collinearity (Tomaschek et al., 2018). To get around this issue, we compare the predictive strength of the abstractness variables using random forest regression, implemented with the ranger package in R (Wright & Ziegler, 2017). Random Forest regression is a machine learning method based on decision trees and recursive partitioning. Each tree is fit to a subset of the data and only uses a subset of the predictors (a third of the available variables in our analyses). The predictions of many hundreds of trees can then be averaged, which helps avoid overfitting and increases accuracy. Because Random Factor regression does not need to estimate a coefficient for every predictor variable in every tree, this approach allows us to avoid issues of collinearity, but the relative importance of two highly correlated predictors can still be compared by assessing the trees in which they do not co-occur.

This approach yields variance importance scores for each of the variables in our model. In Figure XXX, we display the overall ranking of the variance importance scores. We find that XXX is most important, followed by xxx.

Importance plots from random forest regression

We next explore the variability in ranking across languages. intra-class correlation of importance by variable across languages. we find an ICC of R = xxx, indicating xxx stability in relative importance of these predictors across languages.

```{=html}
<!-- how to plot cross-linguistic variability in ordering: maybe horizontal stacked bar graph of importance, with a bar for each language, the importance represented by relative size, and a consistent color for each trait
-->
```
## Multiple behavioral tasks - abstractness effects

picture naming lexical decision

# Discussion

Parents may not have an accurate sense of their children's comprehension of emotion words [@roepstorff2024].
